Order is half of life -- Establishing Totally Ordered HTN Planning Problems is Half the Job (to Getting them Solved More Efficiently)

### 2022-12-01 paper/03--linearization-ground.tex
again some changes

Having presented the approaches for inferring preconditions and effects of compound tasks, we move on to introduce how to linearize a method's task network on top of this information. In particular, we could now view each compound task in the task network to be linearized as an ``action'' (because it is equipped with the inferred precondition and effects just as an action), and thus we could interpret the linearization problem as the problem of finding an executable linearization of a task network consisting solely of actions.

>>>>

We now move on to introduce how to linearize a method's task network. Bearing the inferred preconditions and effects of the compound tasks in a task network, we could formulate this task as follows. Given a task network $tn = (T, \parOrd{}, \alpha)$, we want to find a linearization $\langle t_{1} \cdots t_{|T|} \rangle$ of $tn$ respecting $\parOrd{}$ such that there exists a state sequence $\langle s_{1} \cdots s_{|T| - 1} \rangle$ in which $s_{1} = \effPosExt{\alpha(t_{1})}$, and for each $1 \leq i \leq |T| - 1$, $\preCondExt{t_{i + 1}} \subseteq s_{i}$ and $s_{i} \to_{\alpha(t_{i + 1})} s_{i + 1}$, namely we can view each compound task as an ``action'' and interpret linearizing a task network as the problem of finding an \emph{executable} linearization of the task network consisting solely of actions.

This task is \NP{}-complete, as proved by \shortCite{Erol1996HTNComplexity}. Thus, under the conjecture that $\PTIME{} \neq \NP{}$, no polynomial time algorithms exist that can solve this problem precisely. In order to avoid high computational cost, we present here a poly-time procedure which approximates a linearization which is \emph{most likely} to be executable.

One essential observation for making such an approximation is that in any executable linearization, the precondition of every task except the first one must be satisfied, i.e., in any such linearizations, the following criteria must be satisfied: 
\begin{compactenum}[1)]
    \item if $t$ is a task that is not at the beginning of the sequence, and there is a proposition $p \in \preCondExt{t}$, then there must be a task $t'$ with $p \in \effPosExt{t'}$ placed before $t$ (because otherwise $t$'s precondition cannot be satisfied), and 
    \item for any task $t^{\ast}$ with $p \in \effNegExt{t^{\ast}}$, either
    \begin{compactenum}[a)]
        \item $t^{\ast}$ is placed after $t$, or
        \item $t^{\ast}$ is ordered before some $t'$ with $p \in \effPosExt{t'}$ so that after $t^{\ast}$ deletes $p$, it can be added back by $t'$ and thus satisfies $t$'s precondition. % i.e., there must be a task $t'$ adding back $p$ that is placed somewhere between $t^{\ast}$ and $t$.
    \end{compactenum}
\end{compactenum}

Thus, for linearizing a task network, we would like to add ordering constraints to enforce that the above criteria hold. More concretely, given a task network $tn = (T, \parOrd{}, \alpha)$ with $\parOrd{}^{+}$ being the transitive closure of $\parOrd{}$, we will add an ordering constraint $(t, t')$ with $t, t' \in T$, provided that $(t' ,t ) \notin \parOrd{}^{+}$, if there exists a proposition $p$ such that one of the following conditions holds:
\begin{inparaenum}[1)]
    \item $p \in \preCondExt{t}$ and $p \in \effPosExt{t'}$,
    \item $p \in \effPosExt{t}$ and $p \in \effNegExt{t'}$, or
    \item $p \in \effNegExt{t}$ and $p \in \preCondExt{t'}$. 
\end{inparaenum}
After adding all ordering constraints, we could view the tasks together with the \emph{extended} set of ordering constraints as a graph $G$ where each task is a vertex and each ordering constraint is an edge. A linearization can then be obtained by ordering $G$ in a \emph{topological} order.

Notably, the graph $G$ might \emph{not} have a topological order because it may contain cycles. That is, some ordering constraints we add might conflict with each other. If this is the case, we \emph{randomly} remove some edges $(t, t')$ from $G$ with $(t, t') \notin \parOrd{}^{+}$ (because we \emph{cannot} violate the original ordering constraints in $tn$) until there exist no cycles. We call this step \emph{cycle-breaking}. The entire procedure is shown in Alg.~\ref{alg:Algorithm1}.

% Thus, when finding such a linearization, for every task $t'$ which is not placed at the beginning and every proposition $p \in \preCondExt{t'}$, there is supposed to be another task $t$ with $p \in \effPosExt{t}$ placed before it, i.e., an ordering constraint $(t, t')$ should be added. Similarly, consider a task $t^{\ast}$ with $p \in \effNegExt{t^{\ast}}$ (i.e., $t^{\ast}$ deletes $p$), we would like to place it before $t$ so that after $t^{\ast}$ deletes $p$, it can be added back by $t$ which thus satisfies the precondition of $t'$.
